Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing

### 2019-11-01 Dissonance_Santiago_v1.Rmd
Added discussion based on the files sent by Beatriz, and prepared files for submission

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
  Transportation in most of the world has been dominated for decades by a fascination with the automobile. Nonetheless, there is an increasing recognition that to achieve a variety of economic, environmental, and public health policy goals it is important to attract and retain active travelers and users of public transportation. A challenge with the way transportation policy is developed, however, is that it tends to focus on utilitarian considerations, which may miss other potential aspects of transportation that users value. In particular, subjective wellbeing (SWB) has been proposed as a way to enhance our understanding of the preferences and choices of travellers, as well as a way to evaluate the benefits of transportation beyond utilitarian considerations. The objective of this paper is to analyze the modes that people commonly use, and to what extent they are aligned (or not) with a variety of affective values associated with SWB. In other words, we are interested in the potential for dissonance with respect to the primary mode of travel, from the perspective of affective values. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associate with the affective values of freedom, enjoyment, happiness, poverty, luxury and status. The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, public transportation is most commonly associated with poverty, and the automobil is most commonly associated with luxury and status.

@@

1. Introduction
==========================

@@

A useful lens to investigate the social and psychological functions of transportation is by means of the concept of subjective wellbeing [SWB; see, inter alia, @Devos2013travel; and @Chatterjee2019commuting]. SWB is defined by the OECD as “[g]ood mental states, including all of the various evaluations, positive and negative, that people make of their lives, and the affective reactions of people to their experiences” (**OECD, 2013, p. 10**). As memorably put by Steg [-@Steg2005car]: is the use of car a must, or is it a lust? A close alignment, or consonance, between affective values and the mode of transportation used to travel can result in greater subjective wellbeing and increase the probability of choosing that mode  (i.e., by satisfying the _lust_); on the other hand, dissonance, that is, the lack of concordance between the mode used and affective values would be detrimental to subjective wellbeing, especially of those for whom the use of the mode is a _must_ [@Devos2019satisfaction; @Mokhtarian2018travel]. With these considerations in mind, one aim of the present research is to investigate _who_ experiences dissonance with respect to their primary mode of travel and affective values, namely freedom, enjoyment, happiness, feelings of poverty, luxury, and status. Furthermore, it is possible, even if use of a particular mode is a _must_, that travellers may still _lust_ for something else - in other words, the grass may in fact look greener from the window of the car. For this reason, we also aim to investigate which mode or modes are most commonly associated with affective values, with a focus on those travellers who experience dissonance in their primary mode.

@@

2. Background
============

@@

3. Data and Methods
===================

# 3.1 Context

Words to describe Santiago go here.

# 3.2 Sample

@@

# 4.1 Who experiences dissonance?

@@

```{r table-cross-tabulation-results-without-instrumental, echo = FALSE}
kable(xtab_results,
      "latex",
      digits = 2,
      booktabs = TRUE,
      caption = "\\label{tab:cross-tabulation-results}Percentage of respondents who report mode dissonance with respect to various SWB and instrumental factors",
      escape = FALSE,
      align = c("l", rep("c", 12)),
      col.names = c("Variable", 
                    "Freedom", "$\\chi^2$ p-val", 
                    "Enjoyment", "$\\chi^2$ p-val",
                    "Happiness", "$\\chi^2$ p-val",
                    "Poverty", "$\\chi^2$ p-val",
                    "Luxury", "$\\chi^2$ p-val",
                    "Status", "$\\chi^2$ p-val")) %>%
  collapse_rows(columns = c(3, 5, 7, 9, 11, 13), latex_hline = "major", valign = "middle") %>%
  pack_rows("Age", 1, 3) %>%
  pack_rows("Education", 4, 6) %>%
  pack_rows("Income", 7, 9) %>%
  pack_rows("Mode", 10, 13) %>%
  pack_rows("Commute Time", 14, 17)  %>%
  kable_styling(bootstrap_options = c("striped", "scaled_down")) %>%
  landscape()
```

**FOR DISCUSSION**

@Alayyash2019commute
@Bornioli2019affective
@Devos2019satisfying
@Gatersleben2007affective
@Smith2017commute
@Ye2017satisfaction

@@

The preceding analysis suggests that there is significant mode dissonance along various dimensions and for various affective values. Once we have developed a profile of the travelers who experience mode dissonance, we are interested in what modes they tend to associate with various affective values. For this analysis, we employ a visualization technique known as a mosaic plot that is used for exploring categorical data [@Friendly1994mosaic; @Hofmann2000exploring]. A mosaic plot is a visual representation of a multi-way contingency table, whereby the dimensions of each tile in the mosaic are proportional to the values in the underlying table.

We begin by plotting the cross-tabulation of primary mode of travel to mode associated with each affective value (see Figure \ref{fig:mosaic-plots-by-attribute}). The height of tile $j$ in these mosaics corresponds to the proportion of travellers who primarily travel by mode $j$, hence we can see that the most common primary mode of travel in this sample is by public transportation, followed by car, active travel, and other (also see Figure \ref{fig:primary-mode-travel}). Notice that there are some small discrepancies in the heights across affective values: these are caused by a small number of non-responses. Next, the width of the tile $k$ in the mosaics is proportional to the frequency with which a mode was named in relation to affective value $i$. Recall that respondents could nominate more than one mode in response to each affective value, and some did. For this reason, we used sampling weights, so that the total number of responses corresponds to the size of the sample.

In terms of Freedom, we see that Car users frequently identify this affect with the use of Car; however, those who do not, very often select Active Travel as the mode that better represents Freedom. A majority of active travellers, on the other hand, identify Active Travel as the mode associated with Freedom; few of them attach this value to Car, and even fewer to public transportation and other modes. Travellers whose primary mode of travel is public transportation seldom associate this mode with Freedom, and in fact more frequently relate this affective response to Active Travel followed by Car. Users of other modes seem to be split almost equally in their attribution of Freedom, between Active Travel, Car, and Other. In general, Active Travel is the mode most often identified as embodying Freedom, with the exception of Car travellers.

@@

Three other affective responses show a marked difference. Public Transportation was seldom associated by respondents to feelings of Freedom, Enjoyment, and Happiness; on the other hand, it is the mode that is most commonly selected by respondents for evoking feelings of Poverty - even by users of Publi Transportation. Active Travel was the second most common response for its association with Poverty. Likewise, Whereas Active Travel was often associated to feelings of Freedom, Enjoyment, and Happiness, the mode most frequently associated with Luxury and Status is the Car, although active travellers seem to be somewhat more resistant to this association compared to respondents who travel by Car, Public Transportation, and Other modes.

@@

>>> Efectivamente lujo y estatus parecen similares a primera vista (Figura 1). En general, estos dos valores afectivos son asociados mucho mas comunmente con el automovil. Por otro lado, hay algunas diferencias interesantes cuando se les ve desde la perspectiva de diferentes viajeros.

>>> Por ejemplo, la tendencia a asociar estatus con el automovil tiende a incrementarse con la edad, con la posible excepcion de los usuarios de transporte publico, para quienes otras opciones (posiblemente taxi), ganan en estatus para los viajeros mayores (>54) a costa de otros modos, incluido el automovil. En contraste, el automovil pierde en el valor de lujo con la edad, mientras que otras opciones pasan a ser percibidas como lujo mas comunmente con la edad, incluso para los usuarios de automovil (Figure 2).

>>> se vuelven mas comunmente simbolos de estatus que cuando vemos la manera en que viajeros de diferentes edades proyectan lujo y estatus en varios modos, los usuarios de automovil mayores (>54) muestran menos disonancia en status que en lujo, mientras que los viajeros activos presentan mas disonancia en estatus que en lujo, y de hecho proyectan estatus mas comunmente en el automovil.

>>> La manera en que el lujo se asocia de manera relativamente mas frecuente con otras opciones que el estatus tambien se observa cuando vemos los diferentes niveles de educacion. La excepcion parece ser los usuarios de transporte activo, especialmente  aquellos con educacion superior y de post-grado, que asocian tanto estatus como lujo con el transporte activo mas que los usuarios de otros modos.

@@

Words go here to discuss results by age.

@@

Words go here to discuss results by education.

```{r figure-mosaic-plots-by-attribute-and-education, echo=FALSE, cache = TRUE,  fig.cap="\\label{fig:mosaic-plots-by-education}Mosaic plots for affective values by education; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value (p-values are for Cochran-Mantel-Haenszel Chi-Squared Test)"}

@@

Words go here to discuss results by income.

```{r figure-mosaic-plots-by-attribute-and-income, echo=FALSE, cache = TRUE,  fig.cap="\\label{fig:mosaic-plots-by-income}Mosaic plots for affective values by income; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value (p-values are for Cochran-Mantel-Haenszel Chi-Squared Test)"}

@@

Words go here to discuss results by time.

@@

***FOR DISCUSSION***

@Devos2018people

@@

Accordingly, the aim of this paper is to analyse some of the factors that can potentially affect the levels of SWB in the wide spectrum of public, private and active transport modes in the context of Santiago de Chile. Previous research suggests that public transport users experience a negative gap in desired and actual travel time, whereas active travellers don’t seem to mind somewhat longer trips than what strictly needed (e.g., Páez and Whalen, 2010). In the context of a Latin American country, historically the poor travel experience of public transit users and active travellers reflects the social inequalities of social groups that have no alternative but to move in these types of modes while living far from the work centres and main activities. The paper aims to reflect on the possibilities of increasing the SWB while travelling, and how this could increase the attractiveness of public transport and active travel in the socio-demographic groups studied. This can help not only to reduce the use of private transport but also decrease the inequality gap between these users.

The paper explores how positive affective and instrumental/utilitarian factors of SWB correspond to the use of the primary mode of transportation. It uses an analysis of correspondence that considers the actual primary mode versus the mode that individuals associated to both utilitarian (designed to be useful or practical rather than attractive) and affective (expressing a person's feelings) SWB factors. For example, it will be analysed if a person who uses public transport as the primary mode, tends to associate these modes with positive affective factors (as freedom, enjoyment or happiness) or positive utilitarian factors (as security, time-savings or efficiency). We hypothesize that there is a more persistent non-correspondence or mismatch within the more disadvantage groups or within the users of public transport.

>>>>

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
  Subjective wellbeing is a topic that has attracted considerable attention in recent years due to the way it correlates with health. From a transportation perspective, there is a burgeoning literature on the way travel can impact subjective wellbeing, and how this, in turn, can influence behavior. After decades of fascination with the automobile, subjective wellbeing is increasingly recognized as an alternative approach to assess transportation, and to gain valuable insights to attract and retain active travelers and users of public transportation. The objective of this paper is to analyze a number of affective values associated with subjective wellbeing and the modes of transportation that people commonly use. In particular, we are interested in the potential for dissonance with respect to the primary mode of travel. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associate with the affective values of freedom, enjoyment, happiness, poverty, luxury and status. The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, public transportation is most commonly associated with poverty, and the automobil is most commonly associated with luxury and status.

@@

1. Introduction
===================

@@

A useful lens to investigate the social and psychological functions of transportation is by means of the concept of subjective wellbeing [SWB; see, inter alia, @Devos2013travel; and @Chatterjee2019commuting]. SWB is defined by the OECD as “[g]ood mental states, including all of the various evaluations, positive and negative, that people make of their lives, and the affective reactions of people to their experiences” [@oecd2013oecd, p. 10]. As memorably put by Steg [-@Steg2005car]: is the use of car a must, or is it a lust? A close alignment, or consonance, between affective values and the mode of transportation used to travel can result in greater subjective wellbeing and increase the probability of choosing that mode  (i.e., by satisfying the _lust_); on the other hand, dissonance, that is, the lack of concordance between the mode used and affective values would be detrimental to subjective wellbeing, especially of those for whom the use of the mode is a _must_ [@Devos2019satisfaction; @Mokhtarian2018travel]. With these considerations in mind, one aim of the present research is to investigate _who_ experiences dissonance with respect to their primary mode of travel and affective values, namely freedom, enjoyment, happiness, feelings of poverty, luxury, and status. Furthermore, it is possible, even if use of a particular mode is a _must_, that travellers may still _lust_ for something else - in other words, the grass may in fact look greener from the window of the car. For this reason, we also aim to investigate which mode or modes are most commonly associated with affective values, with a focus on those travellers who experience dissonance in their primary mode.

@@

The structure of the paper is as follows. After these introductory remarks, we follow up with a background section that reviews the literature. Next, we discuss the case study and data used in the research. Then, the analysis and results are presented, before concluding with some discussion and directions for future research.

2. Background
===================

@@

3. Case Study and Data
===================

## 3.1 Context

The case study to contextualize the discussion is Santiago de Chile, the capital of one of the countries with the highest levels of inequality in terms of socio-economic distribution in the world. These inequities have manifested in a disproportionate expense in transport by the different socioe-conomic sectors, a disproportion in travel times and distances travelled. On the other hand, car ownership and use of public transport in the capital are directly conditioned by strong spatial segregation – the higher the income, the higher is the use of the automobile; conversely, and the lower the income, the higher is the predisposition to use public transport. Although transport is a reflection of deep inequities in various dimensions, it has triggered an exceptional discomfort and disagreement with the daily travel experience and quality of life.

## 3.2 Sample

@@

## 4.1 Who experiences dissonance?

@@

```{r table-cross-tabulation-results-without-instrumental, echo = FALSE}
kable(xtab_results,
      "latex",
      digits = 2,
      booktabs = TRUE,
      caption = "\\label{tab:cross-tabulation-results}Percentage of respondents who report mode dissonance with respect to various SWB factors",
      escape = FALSE,
      align = c("l", rep("c", 12)),
      col.names = c("Variable", 
                    "Freedom", "$\\chi^2$ p-val", 
                    "Enjoyment", "$\\chi^2$ p-val",
                    "Happiness", "$\\chi^2$ p-val",
                    "Poverty", "$\\chi^2$ p-val",
                    "Luxury", "$\\chi^2$ p-val",
                    "Status", "$\\chi^2$ p-val")) %>%
  collapse_rows(columns = c(3, 5, 7, 9, 11, 13), latex_hline = "major", valign = "middle") %>%
  pack_rows("Age", 1, 3) %>%
  pack_rows("Education", 4, 6) %>%
  pack_rows("Income", 7, 9) %>%
  pack_rows("Mode", 10, 13) %>%
  pack_rows("Commute Time", 14, 17)  %>%
  kable_styling(bootstrap_options = c("striped", "scaled_down")) %>%
  landscape()
```

@@

The preceding analysis suggests that there is significant mode dissonance along various dimensions and for various affective values. This is for the most part in line with previous research, although by examining different affective values individually instead of a summary measure of wellbeing, we are able to differentiate better in terms of how travellers respond to different affects. Less is known about the values that travellers associate with modes _other_ than the one they use. For this reason, after developing a profile of the travelers who experience mode dissonance in the preceding section, we are now interested in the responses of travellers with respect to the modes they tend to associate with various affective values. For this analysis we employ a visualization technique known as a mosaic plot. Mosaic plots are used in the exploration of multivariate categorical data [@Friendly1994mosaic; @Hofmann2000exploring], and they provide visual representations of multi-way contingency tables. The elements of the mosaic plot are called _tiles_, and the dimensions of each tile are proportional to values of variables in the underlying table.

We begin our analysis by plotting primary mode of travel and the modes associated with each affective value (see Figure \ref{fig:mosaic-plots-by-attribute}). The height of tile $j$ in these mosaics corresponds to the proportion of travellers who primarily travel by mode $j$, hence we can see that the most common primary mode of travel in this sample is by public transportation, followed by car, active travel, and other (also see Figure \ref{fig:primary-mode-travel}). Notice that there are some small discrepancies in the heights across affective values: these are caused by a small number of non-responses. Next, the width of the tile $k$ in the mosaics is proportional to the frequency with which a mode was named in relation to affective value $i$. Recall that respondents could nominate more than one mode in response to each affective value, and some did. For this reason, we used sampling weights, so that the total number of responses corresponds to the size of the sample.

In terms of Freedom, we see that Car users frequently identify this affect with the use of Car; however, those who do not, very often select Active Travel as the mode that better represents Freedom. A majority of active travellers, on the other hand, identify Active Travel as the mode that evokes feelings of Freedom; few of them attach this value to Car, and even fewer to public transportation and other modes. Travellers whose primary mode of travel is public transportation seldom associate this mode with Freedom, and in fact more frequently relate this affective response to Active Travel followed by Car. Users of other modes seem to be split almost equally in their attribution of Freedom, between Active Travel, Car, and Other. In general, Active Travel is the mode most often identified as embodying Freedom, with the exception of Car travellers.

@@

Three other affective responses show a marked difference. Public Transportation was seldom associated by respondents to feelings of Freedom, Enjoyment, and Happiness; on the other hand, it is the mode that is most commonly selected by respondents for evoking feelings of Poverty - even by users of Public Transportation. Active Travel was the second most common response for its association with Poverty. Likewise, Whereas Active Travel was often associated to feelings of Freedom, Enjoyment, and Happiness, the mode most frequently associated with Luxury and Status is the Car, although active travellers seem to be somewhat more resistant to this association compared to respondents who travel by Car, Public Transportation, and Other modes.

@@

@@

There are some interesting differences when viewed from the perspective of travellers of different ages (see Figure \ref{fig:mosaic-plots-by-age}). For example, as seen above, active travel is commonly associated with freedom, happiness, and enjoyment, even by car users, but especially by users of public transportation and other modes. However, when we break this down by age, we notice that this tendency weakens as people age, and older travellers increasingly associate these affective values to the car. Furthermore, the tendency to associate status with the car tends to increase with age, with the possible exception of public transport users, for whom other options (possibly taxi), gain status for older travellers (> 54) at the expense of other modes, including the car. In contrast, the car loses in the value of luxury with age, while other options become more commonly associated with feelings of luxury for older respondents, even for car users (Figure 2). Older car users (> 54) show less dissonance in status than in luxury, while active travellers present more dissonance in status than in luxury - and in fact, they more commonly project status in the car.

@@

Only two affective values were significant in the preceding analysis when seen from the perspective of education, namely luxury and status. As seen in Figure \ref{fig:mosaic-plots-by-education}, the perceptions of luxury and status are fairly consistent when differentiated by education. Most respondents tend to associate luxury and status to the use of car, albeit this association is stronger for status than for luxury. The second mode most often mentioned in relation to luxury is "other" likely due to the flexibility and convenience (but cost) of using taxis. The exception in this case is the traveller with graduate education who uses active modes. This type of respondent, while still more likely to associate luxury and status with the car, more commonly links active travel to these affects than travellers with lower levels of education. Most public transport users tend to associate luxury and status to the car, however, they associate active modes with status more than the same group with respect to luxury.

```{r figure-mosaic-plots-by-attribute-and-education, echo=FALSE, cache = TRUE, fig.width= 6, fig.height=5, fig.cap="\\label{fig:mosaic-plots-by-education}Mosaic plots for affective values by education; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value (p-values are for Cochran-Mantel-Haenszel Chi-Squared Test)"}

@@

Three affective values were significant in the analysis by income; perhaps not surprisingly, all three have an economic dimension. Figure \ref{fig:mosaic-plots-by-income} reflects some important social inequalities associated with the way users feel about transportation in Santiago. When the affective perception of poverty is analyzed, the image shows that generally public transport is the mode most commonly associated with poverty. There are some interesting differences, however, to be observed by income level. Whereas middle income users of public transportation associate this mode to poverty most frequently, this is not the case of high income users of public transportation, who tend to link active travel to poverty with greater frequency. Similarly, active travel is more often associated with poverty by low income travellers than middle or high income travellers. A possible explanation for this is that walking or cycling are related to the idea of a lack of financial means to access other modes of transport. For high income car users, on the other hand, public transportation may have connotations of poverty due to its massive use, whereas active travel may be seen more as a lifestyle alternative.

With respect to luxury and status, we see again that these two values coincide more strongly in their association with the car. Moreover, we also notice that these perceptions become increasingly frequent at higher levels of income. Unlike users of other modes, the high-income public transport user group tends to concentrate a broader association of luxury and status factors with active transport, reinforcing the notion that at higher levels of income, people may see active modes as a lifestyle alternative. High income active travellers, although still more likely to associate luxury and status to the car, are more likely than users of other modes to associate these affects to active travel, perhaps due to an awareness of the benefits of walking and using the bicycle. Furthermore, high-income users of cars are more likely to experience consonance with respect to status and luxury, in what might be a form of self-congratulatory confirmation of success.

```{r figure-mosaic-plots-by-attribute-and-income, echo=FALSE, cache = TRUE, fig.width= 7, fig.height=5, fig.cap="\\label{fig:mosaic-plots-by-income}Mosaic plots for affective values by income; the y-axis is the proportion of respondents by main mode of transportation, and the x-axis is the proportion of modes selected for each value (p-values are for Cochran-Mantel-Haenszel Chi-Squared Test)"}

@@

The last dimension that we examine is typical commute time. Four affective values were significant along this dimension in Table \ref{tab:cross-tabulation-results}. Figure \ref{fig:mosaic-plots-by-travel-time} shows the way travellers associate different affective values to modes of transportation by length of typical commute. The first couple of factors, freedom and enjoyment, are associated with eudaimonic and hedonic factors of wellbeing, as discussed earlier in the paper. On the other hand, the luxury and status are defined mainly by a socio-economic component. As seen in the figure, freedom and enjoyment are most frequently mentioned in relation to active travel, followed by car. However, active travel is more frequently associated with freedom and enjoyment by people who typically undertake relatively short trips (<20 min) or travellers whose typical trips are longer (>60 min). In this respect, active travel may be seen as an ideal mode for trips, and as an enviable mode by people whose trips are longer.

In the case of luxury and status factors, both follow a similar pattern, despite variations in travel times. Despite the travel time exceeding 60 minutes, car users consider that their transport mode is mostly associated with luxury and status. The users of active transport in the sample tend to travel less than 20 minutes, and this is the same segment that considers that the status and luxury factors correspond mainly to the car and then to active transport. Looking at the four factors set out in Figure \ref{fig:mosaic-plots-by-travel-time}, it could be understood that the modes that are associated with intrinsically positive characteristics such as freedom and enjoyment tend to use public transport. On the contrary, factors such as luxury and status tend to have a negative connotation, where the association with the car and other modes of transport (such as the individual taxi and the collective taxi) tends to dominate.

@@

@@

Subjective wellbeing is a topic that has attracted considerable attention in recent years due to its relationship with health. In the field of transportation, there is a large and growing body of literature that documents the way transportation can impact subjective wellbeing. Likewise, there is strong evidence that subjective wellbeing can influence travel behavior. As the world tries to move from a culture dominated by a century-long love affair with the automobile, there is a pressing need to understand how travellers perceive different modes of transportation from the lens of subjective wellbeing. Insights in this regard could prove valuable to develop and implement plans and policies to attract and retain users to healthier, more environmentally friendly transportation options, in particular active travel and public transportation. For this reason, understanding mode dissonance, the extent to which the context of travellers differs from their aspirations, is a worthwhile topic for research.

The premise of the research presented in this paper is that subjective wellbeing is a composite construct of various affective values. In this paper we investigated mode dissonance from the perspective of six affective values associated with subjective wellbeing, namely freedom, enjoyment, happiness, poverty, luxury, and status. The research presented here was based on a sample of travellers in Santiago, the capital of Chile. Participants in this research were asked about their typical mode of travel, and then about the mode or modes that they associate with each of the six affective values. Analysis using hypothesis testing (tests of independence) and vizualization techniques (mosaic plots) uncovered interesting patterns. 
Some of our findings are well aligned with previous research; for example, active travellers experience less dissonance than car users, and users of public transportation experience the most dissonance of all. However, by considering affective values separately instead of aggregating them into a single indicator of subjective wellbeing, we manage to preserve some granularity with respect to various responses. This turns out to be important, because hedonic/eudaimonic values are more frequently related to active travel (freedom, enjoyment, happiness), whereas poverty is more frequently related to public transportation and active travel. Luxury and status, on the other hand, are more frequently associated to luxury and status.

Further delving into the question of which modes are associated with these affective values, we find that there are important differences in terms of the typical mode of travel. Active travellers experience dissonance with relatively little frequently with respect to freedom, enjoyment, and happiness, but when they do, they tend to attach positive values to the car. Car users experience dissonance with respect to these affects more frequently than active travellers, and when they do, they strongly relate possitive hedonic/eudaimonic values to active travel. In other words, it is possible that drivers dream of walking when it comes to feelings especially of enjoyment and happiness. The other side of the coin is also interesting. When it comes to affective values with a stronger socio-economic flavor, such as poverty, luxury, and status, car users tend to experience dissonance less frequently than users of other modes. Active travellers, altough more resistant to the lure of the car compared to users of public transportation, also tend to attach values of luxury and status to the car when they experience dissonance.

An examination of these effects by age, level of education, level of income, and typical trip duration reveals that some of the responses become more pronounced. For instance, older people are less likely than younger people to associate active travel with positive hedonic/eudaimonic affects, and are more likely to attach these values to the car. People with higher incomes are more atuned to the luxury and status values of cars, whereas lower income people are more likely to relate active travel to luxury and status.

Our results not only help to flesh out some ways in which mode dissonance could play out from the perspective of different affects, but also does so in the context of a Latin American country, a region where historically there are deep-seated tabooes with respect to different modes of transportation: the poor travel by public transportation and/or are forced active travellers; where the rich enjoy the luxury of private vehicles and/or are active travellers by choice. In this way, the paper helps us to reflect on the ways the public experiences transportation-related subjective wellbeing, and how this in turn could be used to increase the attractiveness of public transport and active travel in the socio-demographic groups studied.

With respect to opportunities for future research, a possible avenue for extending the findings reported in this paper relates to the use of the modes other than the typical mode of transportation. In a recent paper that investigaed commute satisfaction for car users, Al-Ayyash and Abou-Zeid [-@Alayyash2019commute] considered three models: for current trip satisfaction, remembered satisfaction while using public transport, and current satisfaction using public transport. The findings suggest that low service quality in public transportation can result in a generalised negative perception, and that this perception is more difficult to smooth if commuters do not regularly use public transport. Another avenue for future research could be to consider the mix of modes typically used. While in this paper the analysis focused on the primary mode of transportation, many travellers experience more than one mode of transportation in their daily activities. For this reason, considering the multimodal component of travel would be interesting; for example, future research could consider people who eventually arrive by bicycle to the metro station or people that, after using a colectivo for part of a trip, end their journey by bus.
